---
title: "STAT/MATH 495: Problem Set 04"
author: "WRITE YOUR NAME HERE"
date: "2017-10-03"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE)
set.seed(76)
```

# Collaboration

Please indicate who you collaborated with on this assignment:


# Load packages, data, model formulas

```{r, warning=FALSE}
library(tidyverse)
credit <- read_csv("http://www-bcf.usc.edu/~gareth/ISL/Credit.csv") %>%
  select(-X1) %>%
  mutate(ID = 1:n()) %>% 
  select(ID, Balance, Income, Limit, Rating, Age, Cards, Education)
```

You will train the following 7 models on `credit_train`...

```{r}
model1_formula <- as.formula("Balance ~ 1")
model2_formula <- as.formula("Balance ~ Income")
model3_formula <- as.formula("Balance ~ Income + Limit")
model4_formula <- as.formula("Balance ~ Income + Limit + Rating")
model5_formula <- as.formula("Balance ~ Income + Limit + Rating + Age")
model6_formula <- as.formula("Balance ~ Income + Limit + Rating + Age + Cards")
model7_formula <- as.formula("Balance ~ Income + Limit + Rating + Age + Cards + Education")
```

... where `credit_train` is defined below, along with `credit_test`.

```{r}
set.seed(79)
credit_train <- credit %>% 
  sample_n(20)
credit_test <- credit %>% 
  anti_join(credit_train, by="ID")
```


# RMSE vs number of coefficients

```{r, echo=TRUE, warning=FALSE, message=FALSE}
# Placeholder vectors of length 7. For now, I've filled them with arbitrary 
# values; you will fill these in
RMSE_train <- runif(n=7)
RMSE_test <- runif(n=7)


# Do your work here:


# Save results in a data frame. Note this data frame is in wide format.
results <- data_frame(
  num_coefficients = 1:7,
  RMSE_train,
  RMSE_test
) 

# Some cleaning of results
results <- results %>% 
  # More intuitive names:
  rename(
    `Training data` = RMSE_train,
    `Test data` = RMSE_test
  ) %>% 
  # Convert results data frame to "tidy" data format i.e. long format, so that we
  # can ggplot it
  gather(type, RMSE, -num_coefficients)

ggplot(results, aes(x=num_coefficients, y=RMSE, col=type)) +
  geom_line() + 
  labs(x="# of coefficients", y="RMSE", col="Data used to evaluate \nperformance of fitted model")
```


# Interpret the graph

Compare and contrast the two curves and hypothesize as to the root cause of any differences.



# Bonus

Repeat the whole process, but let `credit_train` be a random sample of size 280
from `credit` instead of 20. Now compare and contrast this graph with the
one above and hypothesize as to the root cause of any differences.

```{r, echo=FALSE}

```
